model_1

### 2020-11-17 DraftReport.Rmd
Added model 1, draft regression table, and conclusion to report

---
title: "Draft Report"
author: "203 Section 6, Team 1"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

- **Common Data across models:**
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(stargazer)
library(tinytex)
```

```{r variables, warning=FALSE}
df <- read.csv("covid-19.csv", header = TRUE)
df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         death_rate_100k = 'Death.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         state_emergency = 'State.of.emergency',
         business_closed = 'Closed.other.non.essential.businesses',
         business_reopen = 'Began.to.reopen.businesses.statewide',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
  select(case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, Black...of.Total.Population, White...of.Total.Population, Percent.living.under.the.federal.poverty.line..2018., Percent.Unemployed..2018., X65.)
# Assign correct data types
cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
df[cols.num] <- sapply(df[cols.num],as.numeric)
df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
head(df)
```

### 1. An Introduction

@@

### 2. A Model Building Process

@@

### Model 1 & Model 2 Building Process

Part of the model aims to investigate the relationship between the covid-19 pandemic status and the state demographic information. The list of demographic variables included in the current dataset is as below.

COVID-19 Cases by Race/Ethnicity:
 - white % of Total Population,
 - Black % of Total Population, 
 - Hispanic % of Total Population, 
 - Other % of Total Population

State Characteristics:
 - Governor,       
 - Population density per square miles, 
 - Population 2018, 
 - Nonelderly Adults Who Have A Pre-Existing Condition
 - Percent at risk for serious illness due to COVID	
 - All-cause deaths 2018, 
 - Number Homeless (2019), 
 - Medicaid Expenditures as a Percent of Total State Expenditures by Fund, 
 - Life Expectancy at Birth (years),
 - Percent Unemployed (2018)
 - Percent living under the federal poverty line (2018), 
 - Weekly UI maximum amount with extra stimulus (through July 31, 2020) (dollars), 
 - Median Annual Household Income

Demographics:
 -Children 0-18
 - Adults 19-25
 - Adults 26-34
 - Adults 35-54
 - Adults 55-64
 - 65+

Currently, Case Rate per 100000 is chosen as a dependent variable to represent the pandemic status in each state. Variables as Population density per square miles, Percent Unemployed (2018)/Percent living under the federal poverty line (2018), White % of of Total Population/Black % of Total Population, Percentage of 65+ are chosen as candidate covariates to explore potential connection between pandemic status demographic characteristics, economy status, racial composition.

Through data explorary, most data investigated is clean as numberica type except the black percentage of total population. This information of some states is reported vaguely as less than 0.01, which is replaced as 0 for further work. This could be updated with more precise data with the help of external resources.

Several transformations have been made for variables a Population density per square miles and White % of Total Population.

### Data Explorary
Through data explorary, most data investigated is clean as numberica type except the black percentage of total population. This information of some states is reported vaguely as less than 0.01, which is replaced as 0 for further work. This could be updated with more precise data with the help of external resources.

@@

In addition, correlation values of White% and Black %, Percent Unemployed and Percent living under the federal poverty line shows these two groups of variables are correlated respectively. 
#### 1-a) Varialbe of Case Rate per 100000 is of nearly normal distribution.

```{r}
summary(df$case_rate_100k)

```{r}
ggplot(data = df,
  mapping = aes(x= case_rate_100k))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
  labs(title =  "Histogram of Case rate per 100000", x = "Case per 100000", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 6000, 600))
```

#### 1-a) Varialbe of Population density per squres 
The distribution of the Population density per squares clusters at the two ends, which is caused by the extremely high population density of District of Columbia (11496.81) compared to the 50 states.
While, the logarithm of population density is close to a normal distribution. In addition, an analysis excluding the District of Columbia is another option.

```{r}
summary(df$population_density)
```
```{r}
hist_population_dens <- ggplot(data = df,
  mapping = aes(x= population_density))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,12000,500)) +
  labs(title =  "Histogram of Population density", x = "Population density per square miles", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 12000, 1200))

hist_log_population_dens <- ggplot(data = df, aes(x = log(population_density))) + 
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,10,1)) + 
  labs(title = 'Histogram of Population density', x = 'Log Population density per square miles', y = 'Count')

hist_population_dens/hist_log_population_dens

```
#### 2-a) White Percentage of Total Population/ Black Percentage of Total Population

@@

### Plots
No obvious trend could be easily observed or concluded from the plots regarding the dependent variables and candidate covariates.

@@

### Model Regression

@@

#### Model 1-a
$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 Population.Density.Per.Square.Miles
\end{aligned}
$$

```{r}
model1a <- lm(case_rate_100k ~ population_density, data = df)
coeftest(model1a, vcov = vcovHC)
```
#### Model 1-b
$$
Case.Rate.Per.100000 =\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)
$$
```{r}
model1b <- lm(case_rate_100k ~ log(population_density) , data = df)
coeftest(model1b, vcov = vcovHC)
```

### Model-2 Regression
#### Model 2-a
$$
\begin{aligned}
 Case.Rate.Per.100000 =\beta_0 &+\beta_1 log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total. population)
\end{aligned}
$$

@@

```{r}
model2b <- lm(case_rate_100k ~log(population_density)+ White...of.Total.Population+Percent.living.under.the.federal.poverty.line..2018., data = df)
coeftest(model2b, vcov = vcovHC)
```

$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)^2\\ &+\beta_3(Percen.living.under.the.federal.poverty.line.2018)
\end{aligned}
$$
```{r}
model2b <- lm(case_rate_100k ~log(population_density)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018., data = df)
coeftest(model2b, vcov = vcovHC)
```
$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_3(Percen.living.under.the.federal.poverty.line.2018)
\end{aligned}
$$
```{r}
model2b <- lm(case_rate_100k ~ Percent.living.under.the.federal.poverty.line..2018., data = df)
coeftest(model2b, vcov = vcovHC)
```

@@

$$
\begin{aligned}
Case.Rate.Per.100000=\beta_0 &+\beta_1log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total.population)\\&+\beta_3(Percen.living.under.the.federal.poverty.line.2018)\\&+\beta_5(65+)
\end{aligned}
$$
```{r}
model2d <- lm(case_rate_100k ~log(population_density)+ White...of.Total.Population+Percent.living.under.the.federal.poverty.line..2018.+X65., data = df)
coeftest(model2d, vcov = vcovHC)
```
## Model 2-d
$$
\begin{aligned}
Case.Rate.Per.100000=\beta_0 &+\beta_1log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total.population)^2\\&+\beta_3(Percen.living.under.the.federal.poverty.line.2018)\\&+\beta_4(65+)
\end{aligned}
$$
```{r}
model2d <- lm(case_rate_100k ~log(population_density)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018.+X65., data = df)
coeftest(model2d, vcov = vcovHC)
```

@@

### Model 3 Building Process

    - For Model 3. Policy related variables will be introduced for the analytics. Policy variable "Mandate face mask use by all individuals in public spaces" (renamed as mask_public) is selected as control variable to predict the case rate. Given it has been proved that mask is effective in preventing the transmission of disease, it is expected that the mask policy mandate in public would have an effect in reducing the case rate. 
    There are other variables related to mask such as "No legal enforcement of face mask mandate" and "Mandate face mask use by employees in public-facing businesses", we believe these variables have a lower impact than the policy that is enforced to public. Also, it is very likely that if a state enforce public mask policy, they will by default enforce mask policy for employees in public-facing business. As a result, we believe mask policy for all individuals in public space can be used to effectively represents the policy impact. 
    It is worth to note that, the mask_public is in date format, which has value of zero or a actual date when the policy is enforced. According to the documentation, zero represents "the absence of an order or directive", which can be interpreted as the policy is not enforced by the state explicitly. For a leaner regression, the date values in mask_public column are transformed as value 1, so that we can distinguish whether states has public mask policy or not. It is also noted that by transforming the variable, some important information will be lost because the actual date (early or late) to enforce the policy can also have an impact to the case rate. However, it is hard to measure the time effect as different states that enforced the public mask policy may have different threshold.

@@

```{r}
model3 <- lm(case_rate_100k ~ log(population_density) +white_pct+ mask_public_bool, data = df)
coeftest(model3, vcov = vcovHC)
```

@@

model3 <- lm(case_rate_100k ~ close_period, data = df)
coeftest(model3, vcov = vcovHC)

@@

model3 <- lm(case_rate_100k ~ business_close_diff + mask_public_bool, data = df)
coeftest(model3, vcov = vcovHC)

@@

### 3. Limitations of the Model

The most general requirement for the linear model is that data points be independent and identically distributed (IID). While for the most part the events of one state will primarily affect that single state, travel by individuals spreading the virus across state lines is an inevitable occurence. This may cause the spread of a virus within one state to be influenced by those around it, or generally by its position in the country. While this is acknowledged to occur, the effect is likely minor compared to the demographic and policy data collected state by state which would affect the spread of the virus. Therefore it is relatively safe for this assumption to be met.

Secondly, the process being modeled must be described by a linear conditional expectation function of the variables which are included, in order for the linear regression to be valid. This can be assessed after the fact by visualizing the model over the range of each variable alongside the data, or also by visualizing the residuals over the range of each variable.

Related to the relationship of the residuals to the data, homoskedasticity can be assessed by visualizing the size of residuals across the range of each variable. Alternatively, it can also be done via the Breusch-Pagan test to more objectively determine whether homoskedasticity is not present in the data.

Perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable. Near perfect colinearity is more problematic to detect, but a common hint to its existence is large standard errors on colinear features. Conceptually, the variables which may have an important colinearity in the final model are

Finally, normally distributed errors can be easily assessed by a histogram of the residuals around zero. Finally, normally distributed errors can be easily assessed by a histogram of the residuals around zero.

#### Limitations of Model 2

The I.I.D assumption for CLM is not satisfied perfectly for all the models discussed above, since the states are not isolated from each other. It is impossible that states suffered from pandemic at different levels independently. Actually, the pandemic status across states tends to be highly related to each other.

In addition, all states are involved in the same market economy system even though there are economic performance divisions across the states. Moreover, in most states white race is still dominant in the population. Many states share the similar racial composition.

>>>>

---
title: "Draft Report"
author: "203 Section 6, Team 1"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

 **Initial reassignment of common data used across models:**
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(stargazer)
library(tinytex)
```

```{r variables, warning=FALSE, results='hide'}
df <- read.csv("covid-19.csv", header = TRUE)
df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         death_rate_100k = 'Death.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         state_emergency = 'State.of.emergency',
         business_closed = 'Closed.other.non.essential.businesses',
         business_reopen = 'Began.to.reopen.businesses.statewide',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
         Black...of.Total.Population, White...of.Total.Population,
         Percent.living.under.the.federal.poverty.line..2018., Percent.Unemployed..2018., X65.)
# Assign correct data types
cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
df[cols.num] <- sapply(df[cols.num],as.numeric)
df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
head(df)
```

## 1. An Introduction

@@

## 2. A Model Building Process

@@

### Model Variables

The three models aim to investigate the relationship between the COVID rate, state demographic information, and mask mandates. The list of variables included in the current dataset that pertain to these concepts are listed below.

State Characteristics:

  - Governor
  - Population density per square miles
  - Population 2018
  - Nonelderly Adults Who Have A Pre-Existing Condition
  - Percent at risk for serious illness due to COVID
  - All-cause deaths 2018 
  - Number Homeless (2019)
  - Medicaid Expenditures as a Percent of Total State Expenditures by Fund
  - Life Expectancy at Birth (years)
  - Percent Unemployed (2018)
  - Percent living under the federal poverty line (2018)
  - Weekly UI maximum amount with extra stimulus (through July 31, 2020) (dollars)
  - Median Annual Household Income
  - Closed other non-essential businesses
  - Stay at home/ shelter in place
  - Mandate face mask use by all individuals in public spaces
  - No legal enforcement of face mask mandate

Demographics:

 - Children 0-18
 - Adults 19-25
 - Adults 26-34
 - Adults 35-54
 - Adults 55-64
 - 65+
 - white % of Total Population,
 - Black % of Total Population, 
 - Hispanic % of Total Population, 
 - Other % of Total Population

Currently, Case Rate per 100,000 is chosen as a dependent variable to represent the pandemic status in each state. Variables such as Population density per square miles, Percent Unemployed (2018)/Percent living under the federal poverty line (2018), White % of of Total Population/Black % of Total Population, Percentage of 65+, and mask mandate policies are chosen as candidate covariates to explore potential relationships between pandemic status demographic characteristics, economic status, racial composition, and state policy.

### Data Exploration

Through data exploration, most data investigated is clean as numerical type except the black percentage of total population. This information of some states is reported vaguely as less than 0.01, which is replaced as 0 for further work. This could be updated with more precise data with the help of external resources.

@@

In addition, correlation values of White% and Black %, Percent Unemployed and Percent living under the federal poverty line shows these two groups of variables are correlated respectively.

### Model 1

For the first model, the relationship between the COVID case rate per 100,000 and the population density per square mile of states was analyzed. First, the distribution of COVID case rate per 100,000 dependent variable is examined.

```{r}
summary(df$case_rate_100k)
```

```{r}
ggplot(data = df,
  mapping = aes(x= case_rate_100k))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
  labs(title =  "Histogram of COVID Case Rate per 100,000", x = "Cases per 100,000", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 6000, 600))
```

As can be seen above, the distribution is fairly normal, and given that it has already been standardized as a rate across all states, there is no need to perform any transformations on this variable. Thus, the case rate per 100,000 variable can be leveraged as is as the dependent variable for all three models.

Next, the distribution of population density per square mile variable is examined.

```{r}
histogram_of_pdensity <- df %>%
  ggplot(aes(x = population_density)) + 
  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
  labs(
    title = 'Distribution of Population Density per Square Mile', 
    x = 'Population density per square mile', y = 'Count')

histogram_of_pdensity
```

As can be seen from the histogram, although most of the population density is concentrated in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration. When an analysis is performed, it can be seen that there is only one data sample that is the outlier, which is the District of Columbia (D.C.) with a value of 11,496. This is given due to the fact that D.C. is a district that solely consists of a large city, as mentioned as a possibility in the introduction. Given that this causes the data to be skewed, the logarithm is taken to scale the variable. Alternatively, the data point could have been droppped from the sample, but given the already small sample size, it was determined that a better approach would be to keep it within the sample. Once this transformation was performed, it is shown that there is a relatively normal distribution of population densities (see figure below).

```{r}
# Find the outlier data points
outliers <- subset(df, population_density > 4000)
paste(outliers$State, '=', outliers$population_density)

# Transform the variable by taking the logarithm and assign it to a new variable
df <- df %>% 
  mutate(l_population_density = log(population_density))

# Plot the new distribution in a histogram
histogram_of_pdensity <- df %>%
  ggplot(aes(x = l_population_density)) + 
  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
  labs(
    title = 'Figure 2. Distribution of Population Density per Square Mile without D.C.', 
    x = 'Log of population density per square mile', y = 'Count')

histogram_of_pdensity
```

With the appropriate variables transformed, a plot is created to show the relationship between them.

```{r}
df %>% 
  ggplot(aes(l_population_density, case_rate_100k)) + 
  geom_smooth(se = FALSE) +
  geom_point() +   
  labs(
    title = 'COVID Case Rate due to Population Density', 
    x = 'Log of population density per square mile', 
    y = 'COVID case rate per 100,000'
  )
```

From the above plot, it can be posited that there is no discernible relationship between the variables given the non-linear relationship. In order to test this, the following equation is used to create a regression model to determine the true relationship between the case rate and population density variables.

$$
Case.Rate.Per.100000 =\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)
$$

```{r}
model1 <- lm(case_rate_100k ~ log(population_density) , data = df)
coeftest(model1, vcov = vcovHC)
```

From the results of the regression, two things are identified. Firstly, the result of the coefficient is nowhere near significant, and therefore we fail to reject the null hypothesis that there is no correlation between case rate and population density, and there is no detectable effect of the independent variable population density. However, the coefficient of population density appears to be negative, which would concur with the assumption that the higher the population density, the lower the COVID case rate in a state.

### Model 2

#### 2-a) White Percentage of Total Population/ Black Percentage of Total Population

@@

### Model 2 Plots

No obvious trend could be easily observed or concluded from the plots regarding the dependent variables and candidate covariates.

@@

@@

### Model-2 Regression

#### Model 2-a
$$
\begin{aligned}
 Case.Rate.Per.100000 =\beta_0 &+\beta_1 log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total. population)
\end{aligned}
$$

@@

```{r}
model2bi <- lm(case_rate_100k ~log(population_density)+ white_pct+Percent.living.under.the.federal.poverty.line..2018., data = df)
coeftest(model2bi, vcov = vcovHC)
```

$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)^2\\ &+\beta_3(Percen.living.under.the.federal.poverty.line.2018)
\end{aligned}
$$
```{r}
model2bii <- lm(case_rate_100k ~log(population_density)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018., data = df)
coeftest(model2bii, vcov = vcovHC)
```
$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_3(Percen.living.under.the.federal.poverty.line.2018)
\end{aligned}
$$
```{r}
model2biii <- lm(case_rate_100k ~ Percent.living.under.the.federal.poverty.line..2018., data = df)
coeftest(model2biii, vcov = vcovHC)
```

@@

$$
\begin{aligned}
Case.Rate.Per.100000=\beta_0 &+\beta_1log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total.population)\\&+\beta_3(Percen.living.under.the.federal.poverty.line.2018)\\&+\beta_5(65+)
\end{aligned}
$$
```{r}
model2d <- lm(case_rate_100k ~log(population_density)+ White...of.Total.Population+Percent.living.under.the.federal.poverty.line..2018.+X65., data = df)
coeftest(model2d, vcov = vcovHC)
```

$$
\begin{aligned}
Case.Rate.Per.100000=\beta_0 &+\beta_1log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total.population)^2\\&+\beta_3(Percen.living.under.the.federal.poverty.line.2018)\\&+\beta_4(65+)
\end{aligned}
$$
```{r}
model2d <- lm(case_rate_100k ~log(population_density)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018.+X65., data = df)
coeftest(model2d, vcov = vcovHC)
```

@@

### Model 3

For Model 3, policy related variables will be introduced for the analytics. Policy variable "Mandate face mask use by all individuals in public spaces" (renamed as mask_public) is selected as control variable to predict the case rate. Given it has been proved that mask is effective in preventing the transmission of disease, it is expected that the mask policy mandate in public would have an effect in reducing the case rate. 
    There are other variables related to mask such as "No legal enforcement of face mask mandate" and "Mandate face mask use by employees in public-facing businesses", we believe these variables have a lower impact than the policy that is enforced to public. Also, it is very likely that if a state enforce public mask policy, they will by default enforce mask policy for employees in public-facing business. As a result, we believe mask policy for all individuals in public space can be used to effectively represents the policy impact. 
    It is worth to note that, the mask_public is in date format, which has value of zero or a actual date when the policy is enforced. According to the documentation, zero represents "the absence of an order or directive", which can be interpreted as the policy is not enforced by the state explicitly. For a leaner regression, the date values in mask_public column are transformed as value 1, so that we can distinguish whether states has public mask policy or not. It is also noted that by transforming the variable, some important information will be lost because the actual date (early or late) to enforce the policy can also have an impact to the case rate. However, it is hard to measure the time effect as different states that enforced the public mask policy may have different threshold.

@@

```{r}
model3a <- lm(case_rate_100k ~ log(population_density) + white_pct 
              + Percent.living.under.the.federal.poverty.line..2018.
              + mask_public_bool, data = df)
coeftest(model3a, vcov = vcovHC)
```

@@

model3b <- lm(case_rate_100k ~ close_period, data = df)
coeftest(model3b, vcov = vcovHC)

@@

model3c <- lm(case_rate_100k ~ business_close_diff + mask_public_bool, data = df)
coeftest(model3c, vcov = vcovHC)

@@

## 3. Limitations of the Models

Firstly, the most general requirement for the linear model is that data points be independent and identically distributed (IID). While for the most part the events of one state will primarily affect that single state, travel by individuals spreading the virus across state lines is an inevitable occurence. This may cause the spread of a virus within one state to be influenced by those around it, or generally by its position in the country. In addition in terms of demographics, all states are involved in the same market economy system even though there are economic performance divisions across the states. Moreover, in most states white race is still dominant in the population. Many states share the similar racial composition.

While this is acknowledged to occur, the effect is likely minor compared to the demographic and policy data collected state by state which would affect the spread of the virus. Therefore it is relatively safe for this assumption to be met. 
Secondly, the process being modeled must be described by a linear conditional expectation function of the variables which are included, in order for the linear regression to be valid. This can be assessed after the fact by visualizing the model over the range of each variable alongside the data, or also by visualizing the residuals over the range of each variable. This can be seen across the three models.

Thirdly, related to the relationship of the residuals to the data, homoscedasticity can be assessed by visualizing the size of residuals across the range of each variable. Alternatively, it can also be done via the Breusch-Pagan test to more objectively determine whether homoscedasticity is not present in the data.

Next, perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable. Near perfect colinearity is more problematic to detect, but a common hint to its existence is large standard errors on colinear features. Conceptually, the variables which may have an important colinearity in the final model are variables such as different percentage make ups of race and ethnicity, or related public policy. These variables are appropriately evaluatea and excluded throughout the model building process.

Finally, normally distributed errors can be easily assessed by a histogram of the residuals around zero.

## 4. Regression Table

```{r warning=FALSE}
se.model1 = coeftest(model1, vcov = vcovHC)[ , "Std. Error"]
se.model2 = coeftest(model2bi, vcov = vcovHC)[ , "Std. Error"]
#se.model3 = coeftest(model2b, vcov = vcovHC)[ , "Std. Error"]
#se.model4 = coeftest(model2c, vcov = vcovHC)[ , "Std. Error"]
#se.model5 = coeftest(model2d, vcov = vcovHC)[ , "Std. Error"]
se.model6 = coeftest(model3a, vcov = vcovHC)[ , "Std. Error"]

stargazer(model1, model2bi, model3a, type = "text", omit.stat = "f",
          se = list(se.model6),
          star.cutoffs = c(0.05, 0.01, 0.001), title = "Table 1: The relationship between COVID case rate and population density")

## 5. Conclusion

In conclusion, although there is no evidence to show that although there is a direct correlation between the COVID case rate and population density as originally hypothesized in our research question, it can be seen from further development of the models that there is some correlation in other variables examined, namely the mask mandate variable. As can be seen from model 3, the regression model gives a mask mandate variable coefficient with p-value of 0.003058, which is highly significant. The negative coefficient also suggests that the original conjecture that the implementation of mask policies within a state would decrease the COVID case rate was indeed accurate.

Conversely, the hypotheses that population and demographics would have an effect on the COVID case rate was unable to be accepted. Although upon investigating the poverty line percentage variable in model 2 there was a detection of notable adjacency to a significant result with a coefficient value of 0.06286, it was not enough to be truly significant and reject the null hypothesis. Similarly, the relationship of race/ethnicity as well as age on case rate was also insignificant with no apparent correlation.

Given these results in conjunction with the original aim of the research question, it can be suggested that variables that can be controlled (in this case, mask policy) has a greater effect on stymieing the spread of COVID-19 than the variables that may be inherent to a state, such as population density and age/ethnicity demographics. This is a promising insight, as it provides opportunity for states to implement and apply policies to intervene and control the case rate without being worried about its varying effects depending on population characteristics.

### 2020-11-17 model_1.Rmd
Added model 1, draft regression table, and conclusion to report

---
title: "model_1"
author: "Sam Temlock"
date: "13/11/2020"
output: pdf_document
---

@@

For our first model, we will analyze the relationship between the COVID case rate per 100,000 and the population density per square mile of states. First, we take a look at the distribution of population density per square mile.

@@

As we can see from Figure 1, although most of the population density is concentrated in the 0 to 1500 range, we have some grouping of outliers that are very far from this concentration. When we perform an analysis, we see that there is only one data sample that is the outlier, which is the District of Columbia (D.C.). This is given due to the fact that D.C. is a district that solely consists of a large city, as mentioned as a possibility in the introduction. Given that this causes the data to be skewed, we choose to remove this data point from the sample. Once removed, we see that we now have a relatively normal distribution of population densities (see Figure 2).

@@

# Remove the outlier from the dataset
df_new <- df[ !(df$State %in% outliers$State), ]

# Check to see that D.C. has been removed
paste("Outliers in the new dataframe:", outliers$State %in% df_new$State)

histogram_of_pdensity <- df_new %>%
  ggplot(aes(x = Population.density.per.square.miles)) + 
  geom_histogram(bins = 20) + 
  labs(
    title = 'Figure 2. Distribution of Population Density per Square Mile without D.C.', 
    x = 'Population density per square mile', y = 'Count')

@@

Next, we evaluate the values for case_rate

>>>>

---
title: "model_1"
author: "Sam Temlock"
date: "13/11/2020"
output:
  html_document:
    df_print: paged
---

@@

For the first model, the relationship between the COVID case rate per 100,000 and the population density per square mile of states was analyzed. First, the distribution of population density per square mile variable is examined.

@@

As can be seen from Figure 1, although most of the population density is concentrated in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration. When an analysis is performed, it can be seen that there is only one data sample that is the outlier, which is the District of Columbia (D.C.). This is given due to the fact that D.C. is a district that solely consists of a large city, as mentioned as a possibility in the introduction. Given that this causes the data to be skewed, the logarithm is taken to scale the variable. Once this transformation is performed, it is shown that there is a relatively normal distribution of population densities (see Figure 2).

@@

# Transform the variable by taking the logarithm and assign it to a new variable
df <- df %>% 
  mutate(l_population_density = log(Population.density.per.square.miles))

# Plot the new distribution in a histogram
histogram_of_pdensity <- df %>%
  ggplot(aes(x = l_population_density)) + 
  geom_histogram(bins = 20) + 
  labs(
    title = 'Figure 2. Distribution of Population Density per Square Mile without D.C.', 
    x = 'Log of population density per square mile', y = 'Count')

@@

Next, the distribution for the case rate of the states is examined for normality.

```{r}
histogram_of_case_rate <- df %>%
  ggplot(aes(x = Case.Rate.per.100000)) + 
  geom_histogram(bins = 20) + 
  labs(
    title = 'Figure 3. Distribution of Case Rate per 100,000', 
    x = 'Case rate per 100,000', y = 'Count')

histogram_of_case_rate
```

As can be seen above in Figure 3, the distribution is fairly normal, and given that it has already been standardized as a rate across all states, there is no need to perform any transformations on this variable.
